### 2025-04-19 thesis/thesis.tex
Methodology

\subsubsection*{RQ1 — Are natural language queries performant and accurate enough to find application in the real world?}

The primary research questions when it comes to natural language database interfaces evolves around their
semantic accuracy and reliability. LLMs have notoriously been known for their ability to hallucinate /
produce false, but promising outputs. This behaviour can be especially  dangerous when opting for data
driven decisions that rely on false data due to a mistranslation from natural language to SQL. LLMs could
cause hard to understand/debug behaviour, like false computation of distributions when the intermediate
format is not being shown to the user. This thesis tries to determine whether such hallucinations could be
reasonably prevented (ie. through pre and post-processing if LLM input/output).

Additionally, this research eveluates the associated performance and hardware requirements, as
a practical deployment of such a natural language database interface outside of research situations
is guarded by the ability to:

\begin{enumerate}
    \item Run on reasonable, mass available hardware  (e.g. excluding high end research GPUs).
    \item Run within a suitable timeframe that is not causing a significant performance degredation of the 
          overall system / delay to the end user.
\end{enumerate}

@@

\subsubsection*{RQ3 — What are possible approaches for increasing semantic accuracy of queries?}

@@

\subsection{Methodological Approach}
\subsection{Structure of the Thesis}

@@

% Theoretical Foundations
\section{Theoretical Foundations}

\subsection{Generative Pretrained Transformers (GPT)}
\subsubsection{Architecture and Functionality}
\subsubsection{Training and Fine-tuning}
\subsubsection{Application Areas}

\subsection{PostgreSQL as a Database System}
\subsubsection{Architecture of PostgreSQL}
\subsubsection{Extension Capabilities}
\subsubsection{PostGIS and Other Extensions as Examples}

\subsection{Embedding AI Models in Database Systems}
\subsubsection{Current Approaches and Solutions}
\subsubsection{Technical Challenges}
\subsubsection{Benefits of Integration}

>>>>

\subsubsection*{RQ1 — Are natural language database interfaces feasible for real world application?}

The primary research questions when it comes to natural language database interfaces evolve around their
semantic accuracy and reliability, therefore questioning their feasibility for real world usage.
LLMs have notoriously been known for their ability to hallucinate / produce false, but promising outputs.
This behaviour can be especially  dangerous when opting for data driven decisions that rely on false data
due to a mistranslation from natural language to SQL. LLMs could cause hard to understand and debug behaviour,
like false computation of distributions when the intermediate format is not being shown to the user. This
thesis tries to determine whether such hallucinations could be reasonably prevented and whether the associated
performance and hardware requirements are suitable for a real world deployment, outside of research situations.

Specifically the two big underlying questions are:

\begin{enumerate}
    \item Is the semantic accuracy of natural language database interfaces high enough to yield a noticable
          benefit to users?
    \item Is it possible to run such an interface on reasonable, mass available hardware (e.g. excluding high end research GPUs).
\end{enumerate}

@@

\subsubsection*{RQ3 — What are effective approaches for increasing semantic accuracy of queries?}

@@

@@

\subsection{Methodological Approach}

This thesis is following a research and development methodology in order to implement a natural language
interface for databases, in particular postgres is used.

\begin{enumerate}
    \item \textbf{Literature Review} — An analysis of the existing research in the fields of
          natural language interfaces (NLI) for databases, GPU integration for acceleration
          of database operations, and LLM/AI Model integration within database systems.
          This phase establishes the theoretical foundation for this research and identifies current 
          state-of-the-art approaches, their benefits and shortcomings.
     \item \textbf{System Design} — Design of a system architecture that can utilize GPU acceleration
          for LLM integration from within postgres. The primary goals of the system design phase
          are to arrive at an architecture that yields low latency natural language processing,
          schema-aware SQL query generation, ambiguity detection and resolution whilst maintaining
          a high semantic accuracy.
     \item \textbf{Implementation} — The implementation of a PostgreSQL extension according to the
          above system design that relies on \texttt{rust} and \texttt{pgrx}. This extension will
          provide a GPU accelerated framework for executing LLMs, implement a natural language
          to query generation pipeline that relies on the SQL schema and create database functions
          and operators for both query generation and execution.
     \item \textbf{Evaluation and Benchmarking} — An assesment framework and benchmark that introspects
          the implementations performance in multiple dimensions. Namely the most relevant dimensions
          for this thesis are:
          \begin{enumerate}
              \item Semantic Accuracy — Measuring the overall accuracy of results delivered for a given
                    natural language input.
              \item Performance Metric — 
          \end{enumerate}
\end{enumerate}

\subsection{Structure of the Thesis}

% Theoretical Foundations
\section{Related Work}

\subsection{360 Degree vergleichbare paper etc.}
